---
title: "Database Connection and Query Instructions"
author: "Gabriel Singer"
date: "9/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(odbc)
library(dbplyr)
library(DBI)
library(tidyverse)

```

## Connect to the Database
First you need to configure the ODBC connection within Windows.  You will need to setup a User DSN.  It appears that Matt P. needs to allow the connection to occur by messing with the database settings (I have no idea what this actually entails)

```{r connect}
sort(unique(odbcListDrivers()[[1]]))


# set up the connection
con <- dbConnect(odbc(), 
                 Driver   = "SQL Server", 
                 Server   = "169.237.35.195", 
                 Database = "smolt_tracking"
                 )

# view tables in the DB
dbListTables(con)

# explore the felds in a particular table
dbListFields(con, "Detections_all")
dbListFields(con, "Tagged_Fish")


# total number of detections
dets <- tbl(con, "Detections_all")
dets %>% 
  tally()

# total number of tagged fish in the database
n_fish <- tbl(con, "Tagged_Fish")
n_fish %>% 
  tally()

# total number of species
species <- tbl(con, "Tagged_Fish")
species %>% 
  distinct(Species) %>% 
  tally()

```


## Queries
 You can  write SQL querires right in the code chunk in Rmarkdown, you can then output the results fo the query directly into a dataframe that can be used in the Rmarkdown document.

```{sql, connection=con, output.var = "test"}
SELECT "TagID"
FROM "Detections_all"
GROUP BY "TagID"
```

Here is one that will return all of the different study names in the database.  
```{sql, connection=con, output.var = "studyIDs"}
SELECT "StudyID", "Agency"
FROM "StudyID"
```

```{sql, connection=con, output.var = "qry_ws"}
SELECT
	TF.Species, DA.FishID, DA.TagID, TF.StudyID, SI.[PI], 
	TF.[Length], TF.Length_Type, DA.Codespace, TF.DateTagged, TF.Date_Released, 
	TF.Sex, TF.Life_Stage, DA.DetectDate,
	ML.Lat, ML.Lon, DA.Location as DetectionLocation, ML.RiverKm, ML.General_Location, DA.VR2SN, 
	Dep.Depth as DepthOfReceiver, DA.Notes as DeployNotes
FROM
	derived__Detections_all_with_loc_where_known as DA INNER JOIN
	Tagged_Fish as TF ON (DA.FishID=TF.FishID) LEFT OUTER JOIN
	StudyID as SI ON (SI.StudyID=TF.StudyID) LEFT OUTER JOIN
	Monitor_Locations as ML ON (ML.Location=DA.Location) LEFT OUTER JOIN
	Deployments as Dep on Dep.Location=DA.Location and Dep.Start=DA.Start and Dep.Stop=DA.Stop
WHERE
	TF.Species = 'White_Sturgeon' and TF.StudyID in ('WS-UCD-YoloBypass-2012','Yolo_Bypass_WST_2014')
ORDER BY
	DetectDate, FishID, RiverKm, DA.Location

```

query by species and studyID		
```{sql, connection = con, output.var = "dets_by_studyID_and_species"}
	select 
		TF.Species, 
		TF.StudyID, 
		count(DA.DetectDate) as cnt 
	from derived__Detections_all_with_loc_where_known as DA
		left outer join Tagged_Fish as TF on 
			DA.FishID=TF.FishID
	WHERE
	TF.Species = 'Steelhead'
	group by
		TF.Species,
		TF.StudyID
```

```{sql, connection = con, output.var = "CALFED_sth"}
SELECT
	TF.Species, DA.FishID, DA.TagID, TF.StudyID, SI.[PI], 
	TF.[Length], TF.Length_Type, DA.Codespace, TF.DateTagged, TF.Date_Released, 
	TF.Sex, TF.Life_Stage, DA.DetectDate,
	ML.Lat, ML.Lon, DA.Location as DetectionLocation, ML.RiverKm, ML.General_Location, DA.VR2SN, 
	Dep.Depth as DepthOfReceiver, DA.Notes as DeployNotes
FROM
	derived__Detections_all_with_loc_where_known as DA INNER JOIN
	Tagged_Fish as TF ON (DA.FishID=TF.FishID) LEFT OUTER JOIN
	StudyID as SI ON (SI.StudyID=TF.StudyID) LEFT OUTER JOIN
	Monitor_Locations as ML ON (ML.Location=DA.Location) LEFT OUTER JOIN
	Deployments as Dep on Dep.Location=DA.Location and Dep.Start=DA.Start and Dep.Stop=DA.Stop
WHERE
	TF.StudyID in ('CALFED-STH-04','CALFED-STH-05')
ORDER BY
	DetectDate, FishID, RiverKm, DA.Location

```

```{sql, connection = con, output.var = "dets_by_studyid"}
	select 
		TF.Species, 
		TF.StudyID, 
		count(DA.DetectDate) as cnt 
	from derived__Detections_all_with_loc_where_known as DA
		left outer join Tagged_Fish as TF on 
			DA.FishID=TF.FishID
	group by
		TF.Species,
		TF.StudyID


		

```


```{sql, connection = con, output.var = "dets_by_species"}
	select 
		TF.Species, 
		count(DA.DetectDate) as cnt 
	from derived__Detections_all_with_loc_where_known as DA
		left outer join Tagged_Fish as TF on 
			DA.FishID=TF.FishID
	group by
		TF.Species
```



You can also use dbplyr to write sql queries using dplyr verbage
```{r dbplyr_test}
qry_test <- tbl(con, "Tagged_Fish") %>% 
  group_by(StudyID, Species) %>% 
  summarize( N = n(), Weight = mean(Weight), Length = mean(Length, na.rm = T))

show_query(qry_test)

ggplot(qry_test, aes(x = N, y = Length)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

```{r dets_plot}
library(magick)
library(cowplot)

names_clean <- c("Unassigned", "Channel Catfish", "Chinook Salmon", "Green Sturgeon", "Largemouth Bass", "Sacramento Pikeminnow", "Sevengill Shark", "Smallmouth Bass", "Spotted Bass", "Steelhead Trout", "Striped Bass", "White Catfish", "White Sturgeon")



dets_by_species_sort <- dets_by_species %>% 
  mutate(name_cln = names_clean) %>% 
  arrange(cnt)
  
  

sturgeon <- ("images/sturgeon.png")
trout <- ("images/trout-silhouette-flip.png")
lmb <- ("images/lmb.png")
catfish <- ("images/catfish_silhoutte_test.png")
shark <- ("images/shark.png")


p <- ggplot(dets_by_species_sort) + 
  geom_bar(aes(x = reorder(name_cln, cnt), y = cnt/1000000), stat = 'identity') + 
  labs(y = "Number of Detections (Millions)", x = "") + 
  coord_flip() + 
  theme_classic()



ggdraw(p) + 
  draw_image(sturgeon, scale = 0.25, hjust = -0.35, vjust = -0.35) + 
  draw_image(trout, scale = 0.17, hjust = -0.25, vjust = -0.21) +
  draw_image(lmb, scale = 0.17, hjust = -0.15, vjust = -0.03) + 
  draw_image(shark, scale = 0.25, hjust = -0.05, vjust = 0.13) + 
  draw_image(catfish, scale = 0.15, hjust = 0.08, vjust = 0.31) + 
  ggsave("fig_output/detections_classic_unassigned.jpg", dpi = 400, width = 7, height = 4)

  
```

  
  
  
  
  
  